Machine learning enhanced empirical potentials for metals and alloys

### 2020-05-20 doc/papers/ml-eam/manuscript.tex
The revision for PRB
Revision 1 for cpc.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became the 
key approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the first-principle accuracy. In this works, we successfully 
developed a route to express EAM and ADP within machine learning framework in 
highly-vectorizable form and further enhanced the potentials with physical 
constraints. Machine learning can significantly boost their performances with 
few training data. 
For energy and force 
predictions, machine lear̦ned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by our 
potentials are nearly 1000 times faster. Machine learned EAM and ADP can also 
give first-principle accuracy key material properties, such as elastic 
constants, melting temperatures and surface energies. Our results suggest a new 
and systematic route for developing machine learning interatomic potentials. 
\end{abstract}

@@

This paper is organized as follows. Section \ref{sec:method} describes the 
theoretical background of this work, including the formalism of the embedded 
atom method and algorithms of the machine learned EAM. Section \ref{sec:results} 
summarizes the results of ML-EAM and ML-ADP. Further discussions are given in 
section \ref{sec:discussions}.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Transformation
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.3]{figures/transformation.png}
\caption{\label{fig:transformation} Visualization of the descriptor matrix 
$\mathbf{G}'$ of the binary AB system and $\nnl$ is 7. Non-colored cells are 
padded zeros. \textbf{(a)} is a sample $\mathbf{g}_i$ of A-type atom and 
\textbf{(b)} represents a B-type atom.}
\end{figure}

@@

The Rose constraint incoporates the universal equation of state (Rose et al) 
into the total loss function. The Mishin-modified equation 
\begin{equation}
\label{eq:rose_eos}
E(x) = E_{0}\left[
    1 + \alpha x + \beta \alpha^3 x^3 \frac{2x + 3}{(x - 1)^2} \right]
    e^{-\alpha x}
\end{equation}
is used because the original form tends to underestimate energies under high 
pressures. In Equation \ref{eq:rose_eos}, $E_{0}$ is the energy of the 
equilibrium structure, $x = a / a_{0} - 1$ is the relative isotropic scaling 
factor ($a$ is the lattice constant), $\beta$ is a chosen parameter and 
\begin{equation}
\label{eq:rose_alpha}
\alpha = \sqrt{-\frac{9 V_{0} B }{E_{0}}}
\end{equation}
where $V_0$ is the equilibrium volume and $B$ is the corresponding bulk modulus. 
The adoption of the Rose constraint guarantees the exact predictions of bulk 
modulus and the energy-volume curve. The loss of the Rose EOS constraint 
$\mathbf{L}^{\mathrm{Rose}}$  is measured as the 2-norm of the energy 
differences between $E(x)^{\mathrm{dft}}$ and their corresponding predicted 
$E(x)$: 
\begin{equation}
\label{eq:rose_loss}
\mathbf{L}^{\mathrm{Rose}} = \sum_{\Omega}{
    \sqrt{\sum_{i}{\left(E(x_i) - E(x_i)^{\mathrm{dft}}\right)^2}}
}
\end{equation}
Here $\Omega$ denotes all selected crystals. In this work, for each included 
crystal, we use the same choices of $x$, which determines the fitting region: 
$x_{i} = x_{0} + \Delta x \cdot i, x_{0} = -0.1, N_{t}^{\mathrm{max}} = 20, 
\Delta x=0.01$. (-10\% to +10\% of the equilibrium volume). In this work, 
$\beta^{\mathrm{Ni}}$ is 0.005 and $\beta^{\mathrm{Mo}}$ is 0.008.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.57]{figures/Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[htp]
\centering
\includegraphics[scale=0.6]{figures/Mo_eam_adp.png}
\caption{\label{fig:Mo_eam_adp} $\rho(r)$, $F(\rho)$, $\phi(r)$ of the original 
Zjw04 EAM, ML-EAM and ML-ADP for elementary Mo. \textbf{(d)} shows the $\mu(r)$ 
and $\omega(r)$ of ML-ADP.}
\end{figure}

@@

% -----------
% Section 2.E Implementation
% -----------
\subsection{Implementation}
\label{sec:implementation}

The implementation of the machine learned EAM and the physical
constraints are beyond the scope of this work. The details will be discussed
in another paper. Here, we will give a brief description.

Both ML-EAM and ML-ADP are implemented within Google's TensorFlow. The 
virtual-atom approach (VAP) is adopted \cite{TensorAlloy} so that we can 
construct a computation graph from atomic positions to total energy directly. 
Thus, atomic forces (Equation \ref{eq:forces}) and virial stress 
(Equation \ref{eq:virial}) can be derived by the AutoGrad module of TensorFlow 
automatically and efficiently. 
\begin{equation}
\label{eq:forces}
\mathbf{F} = -\frac{\partial E^{total}}{\partial \mathbf{R}}
\end{equation}
This direct computation graph also plays a key role in computing analytical 
elastic tensor (Equation \ref{eq:cijkl}) and the Rose loss 
(Equation \ref{eq:rose_eos}).

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3. Results
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Results}
\label{sec:results}

The public Ni-Mo dataset \cite{SNAP} is used to evaluate our approach. This 
dataset is provided by Shyue Ping Ong and co-workers along with their SNAP 
models. It contains 3973 different Ni-Mo solids. All calculations were done by 
VASP \cite{VASP} with the PBE \cite{PBE} functional and the projector 
augmented-wave approach \cite{PAW}.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Elementary Ni and Mo}
\label{sec:elementary}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.65]{figures/Ni_all_eos.png}
\caption{\label{fig:Ni_eos} The energy-volume curves of the original EAM, 
ML-EAM and DFT for the fcc Ni.}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.65]{figures/Mo_all_eos.png}
\caption{\label{fig:Mo_eos} The energy-volume curves of the original EAM, 
ML-EAM, ML-ADP and DFT for the bcc Mo.}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Surface Energy - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.4]{figures/surface_energy.png}
\caption{\label{fig:surface_energy} Comparison of calculated surface energies of 
\textbf{(a)} fcc Ni and \textbf{(b)} bcc Mo.}
\end{figure}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Alloy EOS - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.5]{figures/alloy_eos.png}
\caption{\label{fig:alloy_eos} Energy-volume curves of Ni$_3$Mo and Ni$_4$Mo 
obtained with the original Zjw04 EAM, ML-EAM, ML-ADP, SNAP and DFT.}
\end{figure*}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Mo-Ni}
\label{sec:alloy}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4. Discussions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Discussions}
\label{sec:discussions}

>>>>

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Abstract
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{abstract}
Interatomic potential or force-field plays a vital role in simulating materials. 
Empirical potentials like the embedded atom method (EAM) and its variant 
angular-dependent potential (ADP) have proven successful in modeling many 
metals qualitatively. In the past few years, machine learning has became the 
key approach for modeling interatomic interactions. Powered by big data and 
efficient optimizers, machine learning interatomic potentials can generally 
approximate to the first-principle accuracy. In this works, we successfully 
developed a route to express EAM and ADP within machine learning framework in 
highly-vectorizable form and further enhanced the potentials with physical 
constraints. Machine learning can significantly boost their performances with 
few training data. 
For energy and force 
predictions, machine lear̦ned EAM and ADP can be almost as accurate as the 
spectral neighbor analysis potential (SNAP) on the fcc Ni, bcc Mo and Mo-Ni 
alloy systems, while the molecular dynamics simulations performed by our 
potentials are nearly 1000 times faster. Machine learned EAM and ADP can also 
give first-principle accuracy key material properties, such as elastic 
constants, melting temperatures and surface energies. Our results suggest a new 
and systematic route for developing machine learning interatomic potentials. All 
the new algorithms have been implemented in our program TensorAlloy.
\end{abstract}

@@

This paper is organized as follows. Section \ref{sec:method} describes the 
theoretical background of this work, including the formalism of the embedded 
atom method and algorithms of the machine learned EAM. 
Section \ref{sec:implementation} describes the implementation details.
Section \ref{sec:results} summarizes the results of ML-EAM and ML-ADP. 
Further discussions are given in section \ref{sec:discussions}.

All the new algorithms have been implemented in our program TensorAlloy 
\cite{TensorAlloy}.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Transformation
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.3]{transformation.png}
\caption{\label{fig:transformation} Visualization of the descriptor matrix 
$\mathbf{G}'$ of the binary AB system and $\nnl$ is 7. Non-colored cells are 
padded zeros. \textbf{(a)} is a sample $\mathbf{g}_i$ of A-type atom and 
\textbf{(b)} represents a B-type atom.}
\end{figure}

@@

The Rose constraint incoporates the universal equation of state (Rose et al) 
into the total loss function. The Mishin-modified equation 
\begin{equation}
\label{eq:rose_eos}
E^\mathrm{rose}(x) = E_{0}\left[
    1 + \alpha x + \beta \alpha^3 x^3 \frac{2x + 3}{(x - 1)^2} \right]
    e^{-\alpha x}
\end{equation}
is used because the original form tends to underestimate energies under high 
pressures. In Equation \ref{eq:rose_eos}, $E_{0}$ is the energy of the 
equilibrium structure, $x = a / a_{0} - 1$ is the relative isotropic scaling 
factor ($a$ is the lattice constant), $\beta$ is a chosen parameter and 
\begin{equation}
\label{eq:rose_alpha}
\alpha = \sqrt{-\frac{9 V_{0} B }{E_{0}}}
\end{equation}
where $V_0$ is the equilibrium volume and $B$ is the corresponding bulk modulus. 
The adoption of the Rose constraint guarantees the exact predictions of bulk 
modulus and the energy-volume curve. The loss of the Rose EOS constraint 
$\mathbf{L}^{\mathrm{Rose}}$  is measured as the 2-norm of the energy 
differences between $E^{\mathrm{rose}}(x)$ and their corresponding predicted 
$E(x)$: 
\begin{equation}
\label{eq:rose_loss}
\mathbf{L}^{\mathrm{Rose}} = \sum_{\Omega}{
    \sqrt{\sum_{i}{\left(E(x_i) - E^{\mathrm{rose}}(x_i)\right)^2}}
}
\end{equation}
Here $\Omega$ denotes all selected crystals. In this work, for each included 
crystal, we use the same choices of $x$, which determines the fitting region: 
$x_{i} = x_{0} + \Delta x \cdot i, x_{0} = -0.1, N_{t}^{\mathrm{max}} = 20, 
\Delta x=0.01$. (-10\% to +10\% of the equilibrium volume). In this work, 
$\beta^{\mathrm{Ni}}$ is 0.005 and $\beta^{\mathrm{Mo}}$ is 0.008.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 3. Implementation
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Implementation}
\label{sec:implementation}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.45]{tensoralloy.png}
\caption{\label{fig:tensoralloy} The architecture of TensorAlloy.}
\end{figure}

Fig \ref{fig:tensoralloy} shows the architecture of the TensorAlloy program. The 
program has two major steps: the first step is calculating the losses of energy,
force and stress terms while the second step is measuring the contributions of 
the physical constraints.

For the first step, the implementation is almost the same with our previous 
work \cite{TensorAlloy} except the formula of the potential energy. The direct
computation graph from atomic positions to total energy can be built because of
our virtual atom approach (VAP) \cite{TensorAlloy} and the vectorized EAM theory
\ref{sec:transformation}. Hence, atomic forces and virial stress tensors can be 
calculated automatically thanks to the AutoGrad featue of TensorFlow. The 
batching scheme is adopted in this step.

For the second step, VAP also plays the key role. Only very few specific 
structures are needed to form physical constraints. Thus, the batching scheme is 
no longer a prerequiste. Instead, we use the codes of the 'prediction' phase of 
VAP for flexibility \cite{TensorAlloy}. For each type of physical constraint and 
each selected crystal structure, an independent computation graph shall be 
generated by reusing the model parameters.

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EAM - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.57]{Ni_eam.png}
\caption{\label{fig:Ni_eam} $\rho(r)$, $F(\rho)$ and $\phi(r)$ of the original 
Zjw04 EAM and the machine learned EAM (ML-EAM).}
\end{figure*}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EAM/ADP - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}[htp]
\centering
\includegraphics[scale=0.6]{Mo_eam_adp.png}
\caption{\label{fig:Mo_eam_adp} $\rho(r)$, $F(\rho)$, $\phi(r)$ of the original 
Zjw04 EAM, ML-EAM and ML-ADP for elementary Mo. \textbf{(d)} shows the $\mu(r)$ 
and $\omega(r)$ of ML-ADP.}
\end{figure}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4. Results
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\section{Results}
\label{sec:results}

The public Ni-Mo dataset \cite{SNAP} is used to evaluate our program. This 
dataset is provided by Shyue Ping Ong and co-workers along with their SNAP 
models. It contains 3973 different Ni-Mo solids. All calculations were done by 
VASP \cite{VASP} with the PBE \cite{PBE} functional and the projector 
augmented-wave approach \cite{PAW}.

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4.A
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Elementary Ni and Mo}
\label{sec:elementary}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Ni - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.65]{Ni_all_eos.png}
\caption{\label{fig:Ni_eos} The energy-volume curves of the original EAM, 
ML-EAM and DFT for the fcc Ni.}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Mo - EOS
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.65]{Mo_all_eos.png}
\caption{\label{fig:Mo_eos} The energy-volume curves of the original EAM, 
ML-EAM, ML-ADP and DFT for the bcc Mo.}
\end{figure}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Surface Energy - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure}
\centering
\includegraphics[scale=0.4]{surface_energy.png}
\caption{\label{fig:surface_energy} Comparison of calculated surface energies of 
\textbf{(a)} fcc Ni and \textbf{(b)} bcc Mo.}
\end{figure}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Figure: Alloy EOS - Plot
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\begin{figure*}
\centering
\includegraphics[scale=0.5]{alloy_eos.png}
\caption{\label{fig:alloy_eos} Energy-volume curves of Ni$_3$Mo and Ni$_4$Mo 
obtained with the original Zjw04 EAM, ML-EAM, ML-ADP, SNAP and DFT.}
\end{figure*}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 4.C
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Mo-Ni}
\label{sec:alloy}

@@

% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
% 
% Section 5. Discussions
%
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\subsection{Discussions}
\label{sec:discussions}
